---
output: github_document
---
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rdlearn

<!-- badges: start -->

<!-- badges: end -->

- Author: Kentaro Kawato, Zhang Yi
- Reference: Zhang Yi, Eli Ben-Michael, and Kosuke Imai. 2023. [“Safe Policy
  Learning under Regression Discontinuity Designs with Multiple Cutoffs.”](http://arxiv.org/abs/2208.13323) 
  arXiv: 2208.13323.
- Package manual: pdf file

## Introduction

The R package `rdlearn` implements the *safe policy learning under regression discontinuity designs with multiple cutoffs* of Zhang et al. (2022). It provides functions to learn improved treatment assignment rules (cutoffs) which are guaranteed to yield no worse overall outcomes than the existing cutoffs.

This document demonstrates how to use the main functions of `rdlearn`. For the replication of the empirical results of Zhang et al., please refer to the vignette.

## Installation

The `rdlearn` package for R can be downloaded using (requires previous installation of the [`remotes` package](https://remotes.r-lib.org/index.html)).

Install the latest release from CRAN:

```{r, eval = FALSE}
remotes::packages("rdlearn")
```

Install the development version from GitHub:

```{r, eval = FALSE}
remotes::install_github("kkawato/rdlearn")
```

Load the package after the installation is complete.

```{r, message=FALSE, warning=FALSE}
library(rdlearn)
```

## Data

We can download the `acces` dataset and apply the proposed methodology to the ACCES (Access with Quality to Higher Education) program, a national-level subsidized loan initiative in Colombia.

The `acces` dataset includes four columns. The `elig` column contains the *outcome* (eligibility for the ACCES program (1: eligible; 0: not eligible). The `saber11` column contains *running variable* (position scores from the SABER 11 exam). The `cutoff` column contains the eligibility *cutoff* for each department, and the `department` column contains the names of the departments.

```{r}
library(rdlearn)

# Load acces data
data(acces)
head(acces)
```

## Main Analysis: Cutoff change relative to the baseline for each department under different smoothness multiplicative factors

First, we demonstrate how to output *a summary of the dataset*, which includes local treatment effect estimates at the baseline cutoffs (such as Table 1 in Zhang et al. (2022)). This can be done as follows:
 
```{r}
rdestimate_result <- rdestimate(
  y = "elig",
  x = "saber11",
  c = "cutoff",
  group_name = "department",
  data = acces
)
print(rdestimate_result)
```

This provides basic information, including the sample size and baseline cutoff for each group, as well as the RD treatment effect and standard error. RD treatment effects marked with an asterisk (*) are significant at the 5% level.

Next, we show how to obtain the safe cutoffs by the proposed algorithm. We use the simulation data B in the Appendix D of Zhang et al. (2022). For the replication of the paper, please refer to the vignette.

Safe cutoffs can be learned as follows:

```{r, warning=FALSE, fig.width = 8, fig.height = 6}
set.seed(1234)
data(simdata_B)
head(simdata_B)

rdlearn_result <- rdlearn(data = simdata_B, y = "Y", x = "X", c = "C", 
                          fold = 2, M = c(0, 1), cost = 0, trace = FALSE)

summary(rdlearn_result)
plot(rdlearn_result, opt = "dif")
```

This plot shows the cutoff changes relative to the baseline for each department under different smoothness multiplicative factors (`M`).

The main function here is `rdlearn`. The arguments include `data` for the *dataset*, `y` for the column name of *outcome*, `x` for the column name of *running variable*, `c` for the column name of *cutoff*, and `group_name` for the column name of *group*. These arguments specify the data we analyze. The `fold` argument specifies *the number of folds for cross-fitting*.

For sensitivity analysis, the `M` argument specifies *the multiplicative smoothness factor* and `cost` specifies *the treatment cost*.

For an `rdlearn_result` object obtained by `rdlearn`, we can use `summary` to display the RD estimates, the obtained safe cutoffs, and the differences between the safe and original cutoffs.

The `plot` function provides a clear visualization of the safe cutoffs. Using `plot(result, opt = "safe")` shows *the obtained safe cutoffs*, while `plot(result,opt = "dif")` shows *the differences between the safe and original cutoffs*.

The `trace` argument can be set to `TRUE` to show progress during the learning process.

## Sensitivity Analysis: Cutoff changes relative to the baseline for each department with varying treatment costs

Next, we implement another *sensitivity analysis*.

In the case of Zhang et al.(2022), we assume the utility function \( u(y, w) = y - C \times w \), where \( y \) is a binary outcome (representing the utility gain from enrollment), \( C \) is a cost parameter ranging from 0 to 1, and \( w \) is a binary treatment indicator (representing the offering of a loan). To explore the trade-off between cost and utility, we conduct a sensitivity analysis for the cost parameter: \( C \).

We use the `sens` function with the `rdlearn_result` object as follows:

```{r, warning=FALSE, fig.width = 8, fig.height = 6}
sens_result <- sens(rdlearn_result, M = 1, cost = c(0, 1), trace = FALSE)
plot(sens_result, opt = "dif")
```

This plot shows the cutoff change relative to the baseline for each department under different values of `cost`.

If the learning process has already been implemented and the `rdlearn_result` object is available, the `sens` function can be used to modify parameters specifically for sensitivity analysis.

## References

Zhang, Y., Ben-Michael, E. and Imai, K. (2022) ‘Safe Policy Learning under Regression Discontinuity Designs with Multiple Cutoffs’, arXiv [stat.ME]. Available at: [http://arxiv.org/abs/2208.13323](http://arxiv.org/abs/2208.13323).

Melguizo, T., F. Sanchez, and T. Velasco (2016). Credit for low-income students and access to and academic performance in higher education in Colombia: A regression discontinuity approach. World Development, 80, 61–77.
