---
title: "Replication by `rdlearn`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(eval = TRUE)
```

## Introduction
The R package `rdlearn` implements the *safe policy learning under regression discontinuity designs with multiple cutoffs* of Zhang et al. (2022). It provides functions to learn treatment assignment rules (cutoffs) which are guaranteed to yield no worse overall outcomes than the existing cutoffs.

In this vignette, we replicate some of the empirical application results from Chapter 6 of Zhang et al. (2022).

## Installation
The `rdlearn` package for R can be downloaded using (requires previous installation of the [`remotes` package](https://remotes.r-lib.org/index.html)).

```{r, eval = FALSE}
remotes::install_github("kkawato/rdlearn")
```

Load the package after completed installation. 
```{r, message=FALSE, warning=FALSE}
library(rdlearn)
```


## Data
For our case study, we download the `acces` dataset and apply the proposed methodology to *the ACCES (Access with Quality to Higher Education) program*, a national-level subsidized loan initiative in Colombia. To qualify for the ACCES program, students must achieve a high score on *the SABER 11*, the national high school exit exam. Each student is assigned a position score ranging from 1 to 1,000, based on each student’s position in terms of 1,000 quantiles.

To select ACCES beneficiaries, the Colombian Institute for Educational Loans and Studies Abroad establishes an *eligibility cutoff* such that students whose position scores are higher than the cutoff become eligible for the ACCES program. The eligibility cutoff is defined separately for each department (or region) of Colombia (see Melguizo et al. (2016) for more details).

The `acces` data set includes four columns. The `elig` column contains *eligibility for the ACCES program *(1: eligible; 0: not eligible). The `saber11` column contains *position scores from the SABER 11 exam*, multiplied by −1 so that the values of the running variable above a cutoff lead to the program eligibility. The `cutoff` column contains the *eligibility cutoff* for each department, and the `department` column contains the names of the departments.

```{r}
library(rdlearn)

# Load acces data
data(acces)
head(acces)
```

## Main Analysis: Cutoff change relative to the baseline for each department under different smoothness multiplicative factors

First, we demonstrate how to output *the summary of Table 1* in Zhang et al. (2022), which includes treatment effect estimates. This can be done as follows:

```{r}
rdestimate_result <- rdestimate(data = acces, y = "elig", x = "saber11", 
                                c = "cutoff", group_name = "department")
print(rdestimate_result)
```

This provides basic information, including the sample size and baseline cutoff for each group, as well as the RD treatment effect and standard error. RD treatment effects marked with an asterisk (*) are significant at the 5% level.

Next, we demonstrate how to obtain *the results of Figure 2*, which shows the cutoff change relative to the baseline for each department under different *smoothness multiplicative factors* (`M`). Due to the computational time required, we are presenting only the first two columns of Figure 2. Learning safe cutoffs can be done as follows:

```{r, warning=FALSE, fig.width = 8, fig.height = 6}
set.seed(12345)

# To replicate exactly, set fold = 20 and M = c(0, 1, 2, 4)
rdlearn_result <- rdlearn(data = acces, y = "elig", x = "saber11", 
                          c = "cutoff", group_name = "department", 
                          fold = 5, M = c(0, 1), cost = 0, trace = FALSE)
summary(rdlearn_result)
plot(rdlearn_result, opt = "dif")
```
This plot shows the cutoff change relative to the baseline for each department under different smoothness multiplicative factors (`M`). 

The main function here is `rdlearn`. The arguments include `data` for the *data set*, `y` for the column name of  *outcome*, `x` for the column name of *running variable*, `c` for the column name of *cutoff*, and `group_name` for the column name of *group*. These arguments specify the data we analyze. 

The `fold` argument specifies *the number of folds for cross-fitting*. 
For sensitivity analysis, the `M` argument specifies *the multiplicative smoothness factor* and `cost` specifies *the treatment cost*. For more detail of `M` 

For the `rdlearn_result` object obtained by `rdlearn`, we can use `summary` to show the RD estimates, the obtained safe cutoffs, and the measures of difference between safe cutoffs and the original cutoffs. 

The `plot` function visually represents safe cutoffs as Figure 2 in Zhang et al.(2022). Setting `plot(result, opt = "safe")` shows *the obtained safe cutoffs*, while `plot(result,opt = "dif")` shows *the differences between the safe and original cutoffs*.

The `trace` argument can be set to `TRUE` when we want to display progress during the learning process.

## Sensitivity Analysis: Cutoff change relative to the baseline for each department with varying cost of treatment

Next, we show results for *another sensitivity analysis*, as shown in Figure 3 in Zhang et al.(2022). For the same reason, we are presenting only some of the columns of Figure 3.

We assume the utility function \( u(y, w) = y - C \times w \), where \( y \) is a binary outcome (representing the utility gain from enrollment), \( C \) is a cost parameter ranging from 0 to 1, and \( w \) is a binary treatment indicator (representing the offering of a loan). To explore the trade-off between cost and utility, we implement sensitivity analysis for \( C \).

We use the `sens` function with the `rdlearn_result` object as follows:

```{r, warning=FALSE, fig.width = 8, fig.height = 6}
# To replicate exactly, set cost = c(0, 0.2, 0.4, 0.6, 0.8, 1)
sens_result <- sens(rdlearn_result, M = 1, cost = c(0, 0.6, 1), trace = FALSE)
plot(sens_result, opt = "dif")
```
This plot shows the cutoff change relative to the baseline for each department under different `cost`. 

If we have already implemented the learning process and have the `rdlearn_result` object, the `sens` function can *modify only the parameters for sensitivity analysis*.

## References
Zhang, Y., Ben-Michael, E. and Imai, K. (2022) ‘Safe Policy Learning under Regression Discontinuity Designs with Multiple Cutoffs’, arXiv [stat.ME]. Available at: [http://arxiv.org/abs/2208.13323](http://arxiv.org/abs/2208.13323).

Melguizo, T., F. Sanchez, and T. Velasco (2016). Credit for low-income students and access to and academic performance in higher education in Colombia: A regression discontinuity approach. World Development, 80, 61–77.
